Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies

### 2021-04-25 ms/sections/discussion2.tex
replaced coastal with beach scrub in main text

Instead, we found differences in the way dewlap color differs between habitats across islands. While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected. On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow). Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than coastal lizards on the former, but a lower UV-reflectance on the latter. Under a sensory-drive scenario, we would have expected the same variables to be consistently divergent between habitats, or at least in a consistent direction.\\

>>>>

Instead, we found differences in the way dewlap color differs between habitats across islands. While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected. On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow). Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than beach scrub lizards on the former, but a lower UV-reflectance on the latter. Under a sensory-drive scenario, we would have expected the same variables to be consistently divergent between habitats, or at least in a consistent direction.\\

### 2021-04-25 ms/sections/methods2.tex
replaced coastal with beach scrub in main text

We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A). These islands were chosen to span the breadth of the West Indian range of \textit{A. sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by its dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{tab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

>>>>

We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A). These islands were chosen to span the breadth of the West Indian range of \textit{A. sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by its dominant vegetation type --- xeric beach scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{tab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

### 2021-04-25 ms/sections/results2.tex
replaced coastal with beach scrub in main text

We tested for variation in \textit{A. sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (coastal scrub, primary coppice and mangroves). We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%). Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was very small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\

@@

On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the coastal habitat were often mistaken for dewlaps from the coppice habitat (Fig. \ref{fig:Abaco}A). Importance analysis revealed that coastal and mangrove lizards mostly differed in reflectance in the UV-end of the spectrum (below 400nm, Fig. \ref{fig:Abaco}B), where mangrove dewlaps had higher UV reflectance relative to coastal lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig. \ref{fig:Abaco}C). Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats, representing a higher UV-reflectance relative to red (Fig. \ref{fig:Abaco}D, E, Table \ref{tab:anova}). Coastal lizards also scored lower on PC3, indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig. \ref{fig:Abaco}D, E). Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig. \ref{fig:Abaco}F, G).\\

On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the coastal habitat (Fig. \ref{fig:Bimini}A). Relatively flat importance profiles for coastal lizards suggested that brightness was used instead of a particular wavelength to identify some of the coastal dewlaps (Fig. \ref{fig:Bimini}B). Indeed, some coastal dewlaps were substantially brighter than the rest (Fig. \ref{fig:Bimini}C), a pattern that was captured by our analysis of variance along PC1 (i.e. brightness, Fig. \ref{fig:Bimini}D, E, Table \ref{tab:anova}). The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig. \ref{fig:Bimini}B), which could reflect the significant difference we detected along PC3 between these two habitats (Fig. \ref{fig:Bimini}D, Table \ref{tab:anova}). Coastal lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig. \ref{fig:Bimini}D, E)), and coastal and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig. \ref{fig:Bimini}D, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig. \ref{fig:Bimini}D, E). On this island, the coastal and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig. \ref{fig:Bimini}F, G).\\

On Cayman Brac, all three habitats could be well discriminated against each other (Fig. \ref{fig:CaymanBrac}A), with UV reflectance appearing to be an important variable differentiating coastal and mangrove dewlaps (Fig. \ref{fig:CaymanBrac}B). In contrast, coppice dewlaps had a relatively flat importance profile, suggesting that brightness made them more distinct rather than any particular wavelength (Fig. \ref{fig:CaymanBrac}B). Consistent with this, coppice dewlaps were significantly different from all other dewlaps along PC1 (Fig. \ref{fig:CaymanBrac}D, Table \ref{tab:anova}). At a distance between 2 and 3km (Fig. \ref{fig:CaymanBrac}F, G), dewlaps in the coastal habitat reflected more red light (as represented by PC2, Fig. \ref{fig:CaymanBrac}D, E) and more UV (as represented by PC3) than in the mangrove habitat. Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig. \ref{fig:CaymanBrac}D, E, Table \ref{tab:anova}).\\

Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors. Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small. Consistent with this, the only significant univariate difference detected was for PC2 between coastal and mangrove lizards, where coastal lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig. \ref{fig:Eleuthera}D, E, Table \ref{tab:anova}).\\

Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between coastal and coppice lizards (Fig. \ref{fig:LittleCayman}A). Mangrove dewlaps were possibly most distinct with respect to their reflectance in short wavelengths (Fig. \ref{fig:LittleCayman}B), with significantly lower UV reflectance (as represented by PC2, Fig. \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}). Coastal lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e. slightly higher UV and red reflectance (as represented by higher PC3 scores), than lizards from the other two habitats (Fig. \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}).\\

On Long Island the three habitats were relatively well discriminated (Fig. \ref{fig:LongIsland}A). Importance profiles indicated that short wavelengths were used to discriminate between coastal and mangrove lizards (Fig. \ref{fig:LongIsland}B). Coastal lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig. \ref{fig:LongIsland}D, E, Table \ref{tab:anova}). Coastal lizards also differed from mangrove lizards along PC4 (Fig. \ref{fig:LongIsland}D), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig. \ref{fig:LongIsland}E). Coppice lizards were significantly darker than mangrove and coastal lizards (PC1, Fig. \ref{fig:LongIsland}D, E, Table \ref{tab:anova}).\\

@@

On South Andros coastal and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig. \ref{fig:SouthAndros}A), with importance profiles supported UV-reflectance as a predictor of coppice lizards (Fig. \ref{fig:SouthAndros}B). Coppice lizards had more curved reflectance profiles than coastal lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig. \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}). Coastal lizards also differed from mangrove lizards in PC4 (Fig. \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).\\

>>>>

We tested for variation in \textit{A. sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves). We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%). Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was very small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\

@@

On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the beach scrub habitat were often mistaken for dewlaps from the coppice habitat (Fig. \ref{fig:Abaco}A). Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the UV-end of the spectrum (below 400nm, Fig. \ref{fig:Abaco}B), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig. \ref{fig:Abaco}C). Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats, representing a higher UV-reflectance relative to red (Fig. \ref{fig:Abaco}D, E, Table \ref{tab:anova}). Beach scrub lizards also scored lower on PC3, indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig. \ref{fig:Abaco}D, E). Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig. \ref{fig:Abaco}F, G).\\

On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the beach scrub habitat (Fig. \ref{fig:Bimini}A). Relatively flat importance profiles for beach scrub lizards suggested that brightness was used instead of a particular wavelength to identify some of the beach scrub dewlaps (Fig. \ref{fig:Bimini}B). Indeed, some beach scrub dewlaps were substantially brighter than the rest (Fig. \ref{fig:Bimini}C), a pattern that was captured by our analysis of variance along PC1 (i.e. brightness, Fig. \ref{fig:Bimini}D, E, Table \ref{tab:anova}). The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig. \ref{fig:Bimini}B), which could reflect the significant difference we detected along PC3 between these two habitats (Fig. \ref{fig:Bimini}D, Table \ref{tab:anova}). Beach scrub lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig. \ref{fig:Bimini}D, E)), and beach scrub and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig. \ref{fig:Bimini}D, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig. \ref{fig:Bimini}D, E). On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig. \ref{fig:Bimini}F, G).\\

On Cayman Brac, all three habitats could be well discriminated against each other (Fig. \ref{fig:CaymanBrac}A), with UV reflectance appearing to be an important variable differentiating beach scrub and mangrove dewlaps (Fig. \ref{fig:CaymanBrac}B). In contrast, coppice dewlaps had a relatively flat importance profile, suggesting that brightness made them more distinct rather than any particular wavelength (Fig. \ref{fig:CaymanBrac}B). Consistent with this, coppice dewlaps were significantly different from all other dewlaps along PC1 (Fig. \ref{fig:CaymanBrac}D, Table \ref{tab:anova}). At a distance between 2 and 3km (Fig. \ref{fig:CaymanBrac}F, G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig. \ref{fig:CaymanBrac}D, E) and more UV (as represented by PC3) than in the mangrove habitat. Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig. \ref{fig:CaymanBrac}D, E, Table \ref{tab:anova}).\\

Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors. Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small. Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig. \ref{fig:Eleuthera}D, E, Table \ref{tab:anova}).\\

Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards (Fig. \ref{fig:LittleCayman}A). Mangrove dewlaps were possibly most distinct with respect to their reflectance in short wavelengths (Fig. \ref{fig:LittleCayman}B), with significantly lower UV reflectance (as represented by PC2, Fig. \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}). Beach scrub lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e. slightly higher UV and red reflectance (as represented by higher PC3 scores), than lizards from the other two habitats (Fig. \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}).\\

On Long Island the three habitats were relatively well discriminated (Fig. \ref{fig:LongIsland}A). Importance profiles indicated that short wavelengths were used to discriminate between beach scrub and mangrove lizards (Fig. \ref{fig:LongIsland}B). Beach scrub lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig. \ref{fig:LongIsland}D, E, Table \ref{tab:anova}). Beach scrub lizards also differed from mangrove lizards along PC4 (Fig. \ref{fig:LongIsland}D), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig. \ref{fig:LongIsland}E). Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig. \ref{fig:LongIsland}D, E, Table \ref{tab:anova}).\\

@@

On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig. \ref{fig:SouthAndros}A), with importance profiles supported UV-reflectance as a predictor of coppice lizards (Fig. \ref{fig:SouthAndros}B). Coppice lizards had more curved reflectance profiles than beach scrub lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig. \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}). Beach scrub lizards also differed from mangrove lizards in PC4 (Fig. \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).\\
